The \dealii Library, Version 9.7

### 2025-07-28 9.7/paper.tex
add taskflow section
Make some progress.

The major changes of this release are:
%
\begin{itemize}
\item
  \todo[inline]{Update list}
\end{itemize}
%

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
\item
\todo[inline]{Update list}
\end{itemize}
%
The
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_5_2_and_9_6_0.html}{changelog}
\todo{Update}
-- listing more than XXX
\todo{Update}
features and bugfixes --
contains a complete record of all changes; see \cite{changes97}.

@@

@@

The 9.7 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}
     {around 20 incompatible changes};
see \cite{changes97}. Many of these
incompatibilities change internal
interfaces that are not usually used in external
applications. That said, the following are worth mentioning since they
are more broadly visible:
\begin{itemize}
  \item \todo[inline]{Update}

@@

  \item We have continued to unify the various implementations of object
    orientations in the library. While the three booleans orientation, rotation,
    and flip have maintained their original meanings (i.e., they have default
    values of \texttt{true}, \texttt{false}, and \texttt{false}), the default
    value of the combined orientation (available via
    \texttt{TriaAccessor::combined\_face\_orientation()} and
    \texttt{TriaAccessor::line\_orientation()}) is now $0$ rather than $1$. In
    addition, the return type of these functions is now
    \texttt{types::geometric\_orientation}, which is a \texttt{typedef} for an
    8-bit integral type.
\end{itemize}

>>>>

The major changes of this release are:
%
\begin{itemize}
\item
  \todo[inline]{Update list from the subsections below}
\end{itemize}
%

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
\item MPGMRES and GMRES updates
  \todo[inline]{Matthias and Martin, please update (or make this point
  into one or two subsections below).}
  \item FiniteElement::shape\_function\_belongs\_to() 
\todo[inline]{Update list}
\end{itemize}
%
The
\href{https://dealii.org/current/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}{changelog}
-- listing more than 120
features and bugfixes --
contains a complete record of all changes; see \cite{changes97}.

@@

\subsection{Support for simplex and mixed meshes}

\todo[inline]{Perhaps talk about this point:
  GridGenerator::convert\_hypercube\_to\_simplex\_mesh() can now split
  quadrilaterals into two triangles and hexahedra into six
  tetrahedra. These anisotropic splits are useful in contexts where it
  is important to add as few new elements as possible.}

\subsection{Updates of \dealii{}'s interfaces to external libraries}
\label{subsec:external}

\todo[inline]{Perhaps say something about new interfaces to MUMPS and
  PSBLAS, and the updated interfaces to ArborX. We no longer bundle
  TBB and by default build on Taskflow. More interfaces to CGAL. Now
  bundling Kokkos 4.5.1; are we using something new from Kokkos? magic\_enum}

@@

The 9.7 release includes
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_6_0_and_9_7_0.html}
     {around 25 incompatible changes};
see \cite{changes97}. Many of these
incompatibilities remove previously deprecated functions or classes,
or require now widely available but newer versions of external
dependencies than previous \dealii{} versions. Others
change internal
interfaces that are not usually used in external
applications. The following are worth mentioning since they
are more broadly visible:
\begin{itemize}
  \item \todo[inline]{Update}

@@

  \item We have continued to unify the various implementations of object
    orientations in the library. While the three booleans orientation, rotation,
    and flip have maintained their original meanings (i.e., they have default
    values of \texttt{true}, \texttt{false}, and \texttt{false}), the default
    value of the combined orientation (available via
    \texttt{TriaAccessor::combined\_face\_orientation()} and
    \texttt{TriaAccessor::line\_orientation()}) is now $0$ rather than $1$. In
    addition, the return type of these functions is now
    \texttt{types::geometric\_orientation}, which is a \texttt{typedef} for an
    8-bit integral type.
  \todo[inline]{I didn't pay attention how the entry above came into
    this paper. David: Is this an incompatibility for the current
    release? If so, please just delete this note. Otherwise, please
    remove the bullet point and this note.}
  \item The \step{52} tutorial has been removed. This program used to
    show the use of time stepping functionality in \dealii{}, but
    this functionality was rudimentary compared to what packages such
    as SUNDIALS or PETSc TS offer. As a consequence, \step{52} did
    not show the advanced techniques we would like to promote, and it
    is now superseded by \step{86} (which builds on PETSc TS).
  \item The \texttt{parallel::distributed::SolutionTransfer} class has
    been merged with the \texttt{SolutionTransfer} class, with the
    former now being deprecated. In the process of cleaning up the
    interface, we have also removed some rarely used member functions
    of the latter that prevented us from unifying the two classes.
  \item The process of removing cycles between header files for the
    modules project (see Section~\ref{subsec:modules}) has led to a
    small number of cases where a \dealii{} header file was
    transitively included into a user program, but now no longer
    is. This is easily fixed by adding an explicit \texttt{\#include}
    statement to the user program for the file that is necessary to
    access declarations the compiler now no longer sees.
  \item The \texttt{MatrixOut} class, used to create graphical
    representations of matrices, now defaults to creating sparse
    output, only showing nonzero entries. This vastly increases the
    size of matrices that can be visualized.
  \item The \texttt{FiniteElement} class has a number of functions
    that allow querying properties of implementations in derived
    classes. These include asking whether a derived class is able to provide
    ``(generalized) support points'', i.e., whether the element
    defines its shape functions via nodal interpolation (or via
    quadrature-based integrals) and can return an array of these
    points. These functions returned \texttt{false} for the
    \texttt{FE\_Nothing} element -- an element that has no degrees of
    freedom and consequently no shape functions. This answer was not
    wrong, but missed the point: Codes using these functions want to
    know whether an element is able to provide arrays with these
    support points, and for \texttt{FE\_Nothing} the answer should be
    ``yes'' (i.e., \texttt{true}): The class \textit{can} provide such
    an array, it will just be empty.
\end{itemize}
